---
title: "climate_data_analysis"
author: "Anna Shipunova"
date: "12/5/2020"
output: html_document
---

```{r setup}
if(!file.exists("./storm_data")){ dir.create("./storm_data") }
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "./storm_data")
library(readr)

```

## Title:   
Your document should have a title that briefly summarizes your data analysis

## Synopsis:  
Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences.

## Data Processing  
There should be a section titled Data Processing which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the \color{red}{\verb|cache = TRUE|}cache = TRUE option for certain code chunks.

## Results  
There should be a section titled Results in which your results are presented.

## Questions

1. Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?  
2. Across the United States, which types of events have the greatest economic consequences?


```{r downloading_data, cache=TRUE}
getwd()
file_url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"

file_destination = "StormData.csv.bz2"
download.file(file_url, destfile = file_destination)

storm_data_csv <- read.csv(file_destination)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
