---
title: "Climate data analysis"
author: "Anna Shipunova"
date: "12/5/2020"
output: 
  html_document:
    css: ./storm_data/mystyle.css
---

```{r comment = '', results = 'asis'}

writeLines("
caption {
color: black;
text-align: center;
font-weight: bold;
font-size: x-large;
}
", 
  con = "mystyle.css")
```

```{r setup}
if(!file.exists("./storm_data")){ dir.create("./storm_data") }
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
knitr::opts_knit$set(root.dir = "./storm_data")
library(readr)
library(dplyr)
library(ggplot2)
library(kableExtra)

```

## Which types of severe weather events have most health and economics impact.

## Synopsis:  
Most fatalities are caused by **heat** and most injures by **tornadoes**.
The most property damage caused by "THUNDERSTORM WIND", "FLASH FLOOD" and      "WATERSPOUT".
The most crop damage caused by "DROUGHT".

## Data Processing  
```{r downloading_data, cache = TRUE}
file_url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"

file_destination = "StormData.csv.bz2"
download.file(file_url, destfile = file_destination)

storm_data_csv <- read.csv(file_destination)

```

```{r common_functions}
get_max10 = function(input_data, clmn_name) {
  input_data %>%
    slice_max(order_by = !!as.symbol(clmn_name), n = 10)
}

prepare_for_join = function(input_data, clmn_name) {
  input_data %>% 
    select(EVTYPE, total_number = !!as.symbol(clmn_name)) %>%
    transform(dmg_type = clmn_name) %>%
    group_by(EVTYPE)
}

unique_max = function(input_data) {
  input_data %>% 
    aggregate(total_number ~ EVTYPE, ., max) %>%
    left_join(input_data) 
}

kable_tab = function(input_data, caption = "") {
  input_data %>%
  kable(caption = caption) %>%
  kable_classic(full_width = F) %>%
  kable_material("striped") %>%
  kable_styling(fixed_thead = T) %>%
  scroll_box(height = "20em")
}

by_state = function(input_data, clmn_name) {
  input_data %>% 
  group_by(STATE) %>%
  filter(!!as.symbol(clmn_name) == max(!!as.symbol(clmn_name))) %>%
  select(STATE, EVTYPE) %>% 
  unique()
}

```
<!-- 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?   -->

```{r health}
one_max_fat = storm_data_csv[storm_data_csv$FATALITIES == max(storm_data_csv$FATALITIES),]

one_max_inj = storm_data_csv[storm_data_csv$INJURIES == max(storm_data_csv$INJURIES),]

fat_10_max <- get_max10(storm_data_csv, "FATALITIES")
inj_10_max <- get_max10(storm_data_csv, "INJURIES")

new_data_inj0 <- prepare_for_join(inj_10_max, "INJURIES")
new_data_fat0 <- prepare_for_join(fat_10_max, "FATALITIES")

new_data_fat <- unique_max(new_data_fat0)
new_data_inj <- unique_max(new_data_inj0)

new_data_health <- rbind(new_data_fat, new_data_inj)

```
```{r health by state}

fatal_by_state <- by_state(storm_data_csv, "FATALITIES")
inj_by_state <- by_state(storm_data_csv, "INJURIES")
```

<!-- 2. Across the United States, which types of events have the greatest economic consequences? -->

```{r economics}
max_PROPDMG <- filter(storm_data_csv, PROPDMG == max(PROPDMG)) 
max_CROPDMG <- filter(storm_data_csv, CROPDMG == max(CROPDMG)) 

property_10_max <- get_max10(storm_data_csv, "PROPDMG")
crop_10_max <- get_max10(storm_data_csv, "CROPDMG")

new_data_crop0 <- prepare_for_join(crop_10_max, "CROPDMG")
new_data_property0 <- prepare_for_join(property_10_max, "PROPDMG")

new_data_property <- unique_max(new_data_property0)
new_data_crop <- unique_max(new_data_crop0)

new_data_econ <- rbind(new_data_property, new_data_crop)

```
```{r economics by state}

PROPDMG_by_state <- by_state(storm_data_csv, "PROPDMG")
CROPDMG_by_state <- by_state(storm_data_csv, "CROPDMG")

```

## Results  

### 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?  

<!-- Fatality for all the US: `r one_max_fat$EVTYPE %>% unique()`.   -->
<!-- Injuries for all the US: `r one_max_inj$EVTYPE %>% unique()`.   -->

```{r health_plot}

# ggplot(corr.m, aes(x = reorder(miRNA, -value), y = value, fill = variable)) + 
#   geom_bar(stat = "identity")

new_data_health %>%
ggplot(aes(x = reorder(total_number, EVTYPE), 
           y = total_number, 
           fill = dmg_type)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = total_number),
            hjust = 1.3,
            position = position_dodge(0.9),
            size = 3.5) +
  labs(x = "Event type", 
       y = "Total number", 
       title = "Types of events most harmful with respect to population health") +
  scale_x_discrete(labels = new_data_health$EVTYPE)



# new_data_health %>%
# ggplot(
#   # data = new_data_health, 
#        aes(x = total_number, y = EVTYPE, fill = dmg_type)) +
#   coord_flip() +
#   geom_bar(stat = "identity", position = position_dodge()) +
#   geom_text(aes(label = total_number), 
#             hjust = 1.3,
#             position = position_dodge(0.9), 
#             size = 3.5) +
#   labs(y = "Event type", x = "Total number", title = "Types of events most harmful with respect to population health") + 
#   guides(fill = guide_legend(title = "Harm type"))
```

And the same by state:

```{r comment='', results='asis'}

kable_tab(arrange(fatal_by_state, STATE), caption = "Fatalities by state")

kable_tab(arrange(inj_by_state, STATE), caption = "Injuries by state")

```

### 2. Across the United States, which types of events have the greatest economic consequences?

Property damage: `r max_PROPDMG$EVTYPE %>% unique()`.  
Crop damage: `r max_CROPDMG$EVTYPE %>% unique()`.  

```{r economic_plot}
new_data_econ$total_number <- factor(new_data_econ$total_number, levels = unique(new_data_econ$total_number))

ggplot(data = new_data_econ, 
       aes(x = total_number, y = EVTYPE, fill = dmg_type)) +
   # coord_flip() + 
  geom_bar(stat = "identity", position = position_dodge()) +
  # geom_text(aes(label = total_number), hjust = 1.6,
  #           position = position_dodge(0.9), size = 3.5) +
  labs(x = "Event type", y = "Total number", title = "Types of events most harmful with respect to economics") + 
  guides(fill = guide_legend(title = "Harm type"))


```

The same by state:

```{r comment = '', results = 'asis'}
kable_tab(arrange(PROPDMG_by_state, STATE), 
            caption = "Property damage by state")
 
kable_tab(arrange(CROPDMG_by_state, STATE), 
            caption = "Crop damage by state")
```

