---
title: "CodeBook"
author: "Anna Shipunova"
date: "11/22/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(tidy = FALSE, echo = TRUE)
```

## R Markdown

As it is said in the instructions, input data are collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at [the site where the data was obtained](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones).



## Work description

1. Download input dataset

```{r download, eval=FALSE, tidy=FALSE}
  file_url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  if(!file.exists("./data")){ dir.create("./data") }
  temp <- tempfile()
  download.file(file_url, temp)
```
  
+ To better understand what is in the zip archive we can look inside:
  
  ```{r dir, eval=FALSE}
      list.files(temp)
  ```

# str(temp)
#  chr "/var/folders/j4/90zh63717l98153d_0cyw9140000gn/T//RtmpjTkDQv/file739d5cf0f9e1"

file_destination <- "./data/cleaning_project_data/"

# arch_info <- unzip(temp, list = FALSE)
unzip(temp, exdir = file_destination)
unlink(temp)




```{r cars}
summary(cars)
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
